Crossing Origins by Crossing Formats

### 2013-02-18 main.tex
AFK sleeping

Coercing content to be interpreted
as a different type can be accommodated in the context of the web. 
If the content is loaded using an \emph{image} tag it will be interpreted 
as an image, if a \emph{script} tag is used it will be interpreted as a script. Some 
tags even let the developer decide which type to interpret the content as. 
To prevent abuse, the browsers employ content-type 
sniffing for certain content-types, a practice which has historically led to 
security issues.

@@

Two attacker-centric components require special attention, the infiltration of 
an origin and the exploitation of an infiltrated origin.
%In this situation, the same-origin policy is intended to prevent cross-origin 
%interaction.

@@

In the scenario the attacker infiltrate the origin by uploading a polyglot. Infiltration could, however, be performed in more than one way.
This paper identifies two vectors for infiltrating an origin, via syntax 
injection, or content smuggling.

In a cross-site scripting attack, an HTML document is composed from input 
containing fragments of HTML syntax, causing the execution of attacker supplied 
script code. Similarly, in a syntax injection attack the vulnerable target will 
compose content from attacker controlled input containing syntax of a foreign 
format, thereby creating a polyglot. 
%
The target is assumed to employ server-side cross-site scripting sanitization. Hence, for the attack to be successful, the input must 
bypass this sanitization. The sanitization filter will remove or encode 
problematic characters, based on the context of the web application. In a polyglot attack
the injected format would pose as a new context, unknown to the filter. Chances are that the 
key elements of the injected format will pass the filter unaltered. 
The polyglot content can
be interpreted as either the benign original format or the
malicious attacker-supplied format. 
%

@@

\begin{figure}[h!]
  \center
  \includegraphics[width=0.49\textwidth]{injection-scenario.png}
  \caption{Overview of the syntax injection scenario}
  \label{fig:scenario}
\end{figure}

Previous work documents instances of polyglot attacks based on syntax injection.
Huang et al.~\cite{Huang+:CCS10} describe a cross-origin \emph{cascading style-sheet} (CSS) attack. This 
attack injects fragments of CSS syntax in a HTML 
document, thereby making it a HTML/CSS polyglot. The error-tolerant parsing of 
style sheets allow the polyglot to parsed as valid CSS. The capabilities of CSS 
provide trivial cross-origin leakage. The paper proposes a defense technique which has been
adopted by all major browsers, which implies that the attacks outlined in their paper are now ineffective.
Instead,
Section
\ref{sec:vuln} will show practical attacks based on other formats.

@@

\subsubsection{Content smuggling}
The vulnerable target of a polyglot attack, based on content smuggling, lets 
users to submit files to the service which are subsequently served under the origin of 
the service. Examples of such services are file sharing services, image databases, 
social networks or job broker services. Such a service 
accepts a limited set of file-types, and the files are verified to
be of one of these types, e.g. an image, before being served to the end
user. By submitting a polyglot to such a service, an attacker can trick the
service into serving malicious content under the origin of the service. The
service will verify the file, assuming that it is of the benign format, but once
the content is served an attacker can coerce the content to be interpreted as
the malicious format.

@@

\paragraph{Scenario}

The scenario, illustrated in Figure~\ref{fig:scenario}, describe how the attack proceed, from the infiltration of the origin, to the compromise of the victim.
The victim in this scenario is an authenticated user of a targeted vulnerable website, \emph{vulnerable.com}. 
The target is infiltrated by the attacker (1), manipulating the website to
serve polyglot content. 
The \emph{vulnerable.com} site serves the polyglot content as the benign content type, harmless to the victim. 
At a certain point, while still authenticated to \emph{vulnerable.com}, 
the victim unsuspectingly visits the attackers malicious website, \emph{attacker.com} (2).
The \emph{attacker.com} site embeds the polyglot content from \emph{vulnerable.com} (3), but
coerces the browser to interpret it as the malicious content type (4). 
The capabilities of the malicious format determine the 
severity of the consequences of exploitation (5).

To give more context to the scenario, \emph{vulnerable.com} can be a cloud storage service. 
The attacker infiltrates the target by uploading a polyglot to the service and embed it 
on \emph{attacker.com}. While authenticated to \emph{vulnerable.com}, the unsuspecting victim visits \emph{attacker.com}, the embedded 
polyglot is activated and extracts all the files stored in the victims account.

@@

\begin{figure}[h!]
  \center
  \includegraphics[width=0.49\textwidth]{smuggling-scenario.png}
  \caption{Overview of the content smuggling scenario}
  \label{fig:scenario}
\end{figure}

@@

% dictionary objects
\paragraph{Dictionary objects}
Dictionary objects are a name-value map delimited by the "$<<$" and "$>>$" 
tokens. The names are name objects and the values are objects of any type.
% name objects
Name objects begin with the "/" character, followed by a string of non-whitespace 
characters. Each dictionary has a type, either specified by the "/Type" name or inferred 
from the context in which it occurs. The type declares which kind of element the 
dictionary is describing, e.g. a page or an annotation. Dictionaries form the structure of the document by connecting
objects through references, e.g. relating a page to its contents. %At least one 
%dictionary is required to build a PDF document. 
% actions?
A special type of dictionary describes \emph{actions}. Actions are triggered when a
certain event occur, such as a file is opened or a page is displayed, and the action 
dictionary specify how it should be handled. Actions can be used to, among other 
things, go to a specific page, play a sound, execute JavaScript or launch a command.

% stream objects
\paragraph{Stream objects}
A stream is an unlimited sequence of bytes. A stream object consists of a 
dictionary, describing the stream, and the associated stream delimited by the 
"stream" and "endstream" keywords. PDF supports encoding of streams, in which
case the dictionary describe which filters are required to decode the stream.

@@

% trailer
\subsubsection{Trailer}
The trailer is composed of a trailer dictionary, a pointer to the cross-reference 
section and an end-of-file marker. The trailer dictionary is introduced by the 
"trailer" keyword. It contains among other things a reference to the root of the 
document structure. The "startxref" keyword is followed by the number of bytes from 
the beginning of the file to the first cross-reference section. As with the 
cross-reference section, it can be omitted. The string "\%\%EOF" marks the 
end-of-file, but can be omitted.

\subsection{PDF-based polyglots}
Several design choices in the PDF specification make the format particularly 
suitable for making polyglots. One such design choice is the error tolerant 
parser. This is in part motivated by PDF being a container format, of which the 
implication is that a PDF file can, by design, contain foreign syntax that 
could interfere with the syntax of the file. Another motivation is that PDF 
files are designed to be both forward- and backward-compatible. Readers 
implementing an older version of the specification do not recognize new 
features and behave as if they were not present in the file. 
%
The implementation notes of the specification describes some exceptions to the
requirements of the specification, such as the header can occur 
anywhere within the first 1024 bytes of the file. This flexibility gives plenty of room for 
combining with syntax of another format.
%
The specification declares many components to be required in a
PDF file, but as can be seen in Section~\ref{sec:format}, in practice several 
components can be omitted~\cite{Wolf:CCC10}. The code in Listing~\ref{lst:minimal-pdf} shows the 
minimal syntax required for a malformed, but valid PDF file.
%incrementally 
%updated by appending updated versions of existing objects to 
%the end of the file. In fact, 
%and formats. This and the fact that it is designed to be robust against errors,
%provides us with a format that is powerful, flexible and ideal for creating
%malicious polyglots.

@@

Furthermore the PDF format is of particular interest to us because of its 
many capabilities. Some examples include executing JavaScript, launching commands, 
and issuing HTTP requests. Adobe Reader even includes a flash 
player to play embedded flash files on systems that do
not have a player installed. %Considering that it might be a conscious decision
%of the user to refuse a flash player,

@@

When a PDF document is embedded in a web page, the corresponding plugin 
is executed to render the content. Recall from Section~\ref{sec:cross-origin} 
that the plug-in is selected based 
on the MIME-type, either supplied by the server or in the type attribute, and 
that the browser rely on the plug-in to handle the situation when the MIME-type 
supplied by the server is inappropriate for the plug-in.
In the case of the Adobe Reader plug-in, it disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF.

@@

Based on the PDF samples in Listings~\ref{lst:minimal-pdf}, we can derive 
the set of tokens required to build a PDF. Assuming that alpha-numeric characters 
pass through a sanitization filter unmodified, the set of tokens is $\{\%, <<, >>, /\}$.
As can be noted, there is a small but significant overlap with the tokens of HTML, 
which implies that many websites protected against cross-site scripting attacks 
are also protected against PDF-based polyglot attacks. One of the problems of 
filtering input for inclusion in a web page are the many contexts in which the 
input can be included. A problem made even more complex by the diversity of 
languages the page contains. Language incompatibilities force context dependent 
filtering, where the same input is treated differently based on the context in 
which it is included. In certain contexts angle brackets are often left untouched by filters.

@@

\subsection{Content smuggling}

Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format. Consequently, 
this significantly expands the attack surface, making it important to take 
measures to protect against these attacks.

\paragraph{PDF as the benign format}
Services like job brokers commonly let the user upload a CV in the form of a PDF-file.
Before such PDF-files are published to recruiters, they are verified to not 
contain any malicious payload. An attacker can produce a PDF that is valid and 
benign, but also a polyglot hiding another malicious format, such as Flash. 
A proof-of-concept can easily be created by storing the PDF as a static string 
variable in the malicious flash source code. When the flash is compiled, the 
source of the PDF is part of the output, making it a valid PDF.

@@

\subsection{PDF communication methods}

The embedded document will retain the origin from which it was served. All 
request issued from the document will include any cookies available in the 
browser for the target domain.
When embedded in a web page, many of the PDF communication methods utilize
the underlying browser to make requests by navigating the browser to the target 
URL. This will unload the current page in which the document is embedded. Arbitrary page 
navigation is allowed with respect to SOP, since the response is not visible 
to the document. Such methods can be used in CSRF attacks, but would 
defeat the purpose of requesting information as the document cannot process the 
response.

% Cross-origin leaks
To leak information across origins, a communication channel is 
required between the attacker and the target. PDF provides the necessary means 
to mediate this information leakage. The flow of information can be divided into 
two steps, extracting information from the target and delivering it to the attacker.
Each requiring its own communication methods.

\subsubsection{The target}
There are two methods for a PDF file to communicate with the server it originates from, 
both allowing it to read the response: from JavaScript using XML external entities 
and from an embedded flash file.

\paragraph{XML}
The PDF JavaScript API includes a method to parse an XML document, called 
XMLData.parse. The XML document being processed may rely on external entities, 
which the XML parser will request and include in the document. The request is 
bound to the origin of the document. As the response 
is included in the XML, the result has to be well-formed XML. This puts a 
restriction on the content that can be requested using this method. Considering 
that HTML pages are rarely well-formed this method may be too restricted in practice.
Listing~\ref{lst:xml-pdf} contains an example that uses XML external entities.

\paragraph{Embedded Flash}
The embedded flash file inherits the origin of the document, thus it can 
request and read any document from the originating server. Unlike the XML
method, there is no restrictions on what content can be requested.

\begin{lstlisting}[label=lst:xml-pdf, caption=PDF using XML for communication]
%PDF-
1 0 obj<<>>stream
xml = '<!DOCTYPE foo[<!ENTITY x SYSTEM "'+
	URL+'">]><foo>&x;</foo>'
var doc = XMLData.parse(xml);
app.alert(doc.saveXML())
endstream
trailer<</Root
   <</Pages<<>>
      /OpenAction
      <</S/JavaScript/JS 1 0 R>>
   >>
>>
\end{lstlisting}

\subsubsection{The attacker}
There are a number of ways to communicate with the attacker, either directly 
via the embedding page or through the attackers server. As mentioned previously, 
many methods trigger page navigation in the browser. This will result in a 
noticeable change in the victims browser. While it fulfills the purpose of 
leaking information, an attacker can use more covert methods of communication
to refrain from alerting the victim.

\paragraph{Fragment identifier}
The methods that navigate the browser can still be used for a somewhat covert 
communication with the attacker. Navigating to the same page with a different 
fragment identifier will not reload the page. Embedding the information in the fragment 
identifier of the URL of the embedding website, creates a one-way communication 
channel from the document to the page. Since the fragment identifier is visible 
in the address field of the browser, an alert user would notice it changing.

\paragraph{PostMessage}
The PDF plugin includes an API for communicating directly with the page in 
which it is embedded. The document can send messages to the page using 
the hostContainer.postMessage method. The page registers a handler to receive 
the event and is also able to post messages back to the document.
This two way communication is not visible to the victim.

\paragraph{Embedded Flash}
Cross-origin communication in Flash is restricted by the same-origin policy,
but also support cross-origin communication via  
CORS. Using CORS, a web server can relax SOP to allow access to specified content.
As mentioned, the embedded Flash file inherits the origin of the document.
However, this is not a restriction, since an attacker can set a cross-origin policy 
that allow the request. This allows for a two way communication that is not 
visible to the victim.

@@

\paragraph{Safari and Opera}
The browsers, Safari and Opera, are both susceptible to the attacks outlined as 
per Section~\ref{sec:vuln}, without any restrictions or modifications.

\paragraph{Firefox}
The Mozilla Firefox browser is susceptible to the attacks, with the restriction 
that the PostMessage API is unavailable. This merely prevents one channel of 
communication with the attacker, and does not reduce the threat.

@@

@@

Wolf's OMG WTF PDF presentation~\cite{Wolf:CCC10} is one of the 
inspirations for our work. The presentation explores the liberty
of the PDF format. In addition, it highlights that PDF interpreters
often disregard the specification demands. This particularly
relevant as it allows crossbreeding PDF with such
formats as ZIP and EXE.

>>>>

Coercing content to be interpreted
as a different type can be accommodated in the context of the web. 
If the content is loaded using an \emph{image} tag it will be interpreted 
as an image, if a \emph{script} tag is used it will be interpreted as a script. Some 
tags even let the developer decide which type to interpret the content as. 
To prevent abuse, the browsers employ content-type 
sniffing for certain content-types, a practice which has historically led to 
security issues. 
%

@@

While these articles document known instances of polyglot attacks, the attack 
method has up until now never been generalized. 
\todo{Is this a good place? is it a good way to say it?}
%
Two attacker-centric components require special attention; the infiltration of 
an origin, i.e. the attack vector, and the exploitation of an infiltrated origin, i.e. the payload.
%In this situation, the same-origin policy is intended to prevent cross-origin 
%interaction.

@@

%In the scenario the attacker infiltrate the origin by uploading a polyglot. Infiltration could, however, be performed in more %than one way.
This paper identifies two vectors for infiltrating an origin, either via syntax 
injection, or content smuggling. In the case of syntax injection, existing content is manipulated
to become a polyglot, whereas with content smuggling, a polyglot is used to evade 
content filters.

\subsubsection{Syntax injection}
\label{sec:injection}
In a cross-site scripting attack user input is used to compose an HTML document. 
Fragments of HTML-syntax in the input alters the semantic of the document to 
execute attacker supplied 
script code. Similarly, in a syntax injection attack the vulnerable target will 
compose content from attacker controlled input containing syntax of a foreign 
format, thereby creating a polyglot. The polyglot is benign when viewed as the 
original format, but malicious when viewed as the injected format. While the 
targeted service serves the content as its original format, it is in the hands 
of the attacker to decide how it is interpreted.
%
The targeted service is assumed to employ server-side cross-site scripting 
sanitization. Hence, for the attack to be successful, the input must 
bypass this sanitization. The sanitization filter will remove or encode 
problematic characters related to HTML-syntax, based on the context of the web application. In a polyglot attack
the injected format will pose as a new context, unknown to the filter. Chances are that the 
key elements of the injected format will pass the filter unaltered.

%The polyglot content can
%be interpreted as either the benign original format or the
%malicious attacker-supplied format. 
%

@@

Previous work documents an instance of a polyglot attack based on syntax injection, though not phrased in those terms.
Huang et al.~\cite{Huang+:CCS10} describe a cross-origin \emph{cascading style-sheet} (CSS) attack. This 
attack injects fragments of CSS syntax in a HTML 
document, thereby making it a HTML/CSS polyglot. The error-tolerant parsing of 
style sheets allow the polyglot to parsed as valid CSS. The capabilities of CSS 
provide trivial cross-origin leakage. The paper proposes a defense technique which has been
adopted by all major browsers, which implies that the attacks outlined in their paper are now ineffective.
Instead,
Section
\ref{sec:vuln} will show practical attacks based on other formats.

\paragraph{Scenario}
The scenario, illustrated in Figure~\ref{fig:injection-scenario}, describe how 
the attack proceed, from the infiltration of the origin, to the compromise of the victim.
The victim in this scenario is an authenticated user of the targeted website, \emph{vulnerable.com}. 
The \emph{vulnerable.com} site serves the polyglot content as the benign content type, harmless to the victim. 
At a certain point, while still authenticated to \emph{vulnerable.com}, 
the victim unsuspectingly visits the attackers malicious website, \emph{attacker.com} (1).
The \emph{attacker.com} site serves a page that embeds the content from \emph{vulnerable.com} (2). 
The attacker supplies polyglot syntax in the input to the vulnerable 
service, e.g. in the parameters of the request for the embedded content (3).
The \emph{vulnerable.com} service responds with content of the benign content type (4), but
the way it is embedded in the attackers page coerces the browser to interpret it as the malicious content type.
The malicious payload is executed in the origin of \emph{vulnerable.com} and can 
hence leverage the credentials of the victim to further exploit the vulnerability.

\begin{figure}[t]
  \center
  \includegraphics[width=0.49\textwidth]{injection-scenario.png}
  \caption{Overview of the syntax injection scenario}
  \label{fig:injection-scenario}
\end{figure}

@@

\subsubsection{Content smuggling}
\label{sec:smuggling}
The vulnerable target of a content smuggling attack, lets 
users submit files to the service which are subsequently served under the origin of 
the service. Examples of such potentially vulnerable services are file sharing services, image databases, 
social networks, conference management systems or job broker services. Such a service 
accepts a limited set of file-formats, and the files are run through a filter to verify that they 
belong to this set, e.g. images, before being served to the end
user. The
service will verify the file under the assumption that it conforms strictly to one format. 
By submitting a polyglot to such a service, an attacker can evade the filter as the polyglot does conform to the benign format.
Since the attacker can coerce the content to be interpreted as the malicious format, the
service is tricked into serving malicious content.

%a polyglot assuming that it is of the benign format, but once
%the content is served an attacker can coerce the content to be interpreted as
%the malicious format.

@@

\paragraph{Scenario}
The content smuggling scenario proceeds as illustrated in Figure~\ref{fig:smuggling-scenario}.
The victim in this scenario is an authenticated user of a targeted vulnerable website, \emph{vulnerable.com}. 
The target is infiltrated by the attacker (1), manipulating the website to
serve polyglot content. 
The \emph{vulnerable.com} site serves the polyglot content as the benign content type, harmless to the victim. 
At a certain point, while still authenticated to \emph{vulnerable.com}, 
the victim unsuspectingly visits the attackers malicious website, \emph{attacker.com} (2).
The \emph{attacker.com} site embeds the polyglot content from \emph{vulnerable.com} (3), but
coerces the browser to interpret it as the malicious content type (4). 
The capabilities of the malicious format determine the 
severity of the consequences of exploitation (5).

To give more context to the scenario, \emph{vulnerable.com} can be a cloud storage service. 
The attacker infiltrates the target by uploading a polyglot to the service and embedding it 
on \emph{attacker.com}. While authenticated to \emph{vulnerable.com}, the unsuspecting victim visits \emph{attacker.com}, the embedded 
polyglot is activated and extracts all the files stored in the victims account.

@@

\begin{figure}[t]
  \center
  \includegraphics[width=0.49\textwidth]{smuggling-scenario.png}
  \caption{Overview of the content smuggling scenario}
  \label{fig:smuggling-scenario}
\end{figure}

@@

% dictionary objects
\paragraph{Dictionary objects}
Dictionary objects are a name-value map delimited by the "$<<$" and "$>>$" 
tokens. The names are name objects and the values are objects of any type.
% name objects
Name objects begin with the "/" character, followed by a string of non-whitespace 
characters. Each dictionary has a type, either specified by the "/Type" name or inferred 
from the context in which it occurs. The type declares which kind of element the 
dictionary is describing, e.g. a page or an annotation. Dictionaries form the structure of the document by connecting
objects through references, e.g. relating a page to its contents. %At least one 
%dictionary is required to build a PDF document. 
% actions?
A special type of dictionary describes \emph{actions}. Actions are triggered when a
certain event occur, such as a file is opened or a page is displayed, and the action 
dictionary specify how it is handled. Actions can be used to, among other 
things, go to a specific page, play a sound, execute JavaScript or launch a command.

% stream objects
\paragraph{Stream objects}
A stream is an unlimited sequence of bytes. A stream object shall always be indirect and consists of a 
dictionary, describing the stream, and the associated stream delimited by the 
"stream" and "endstream" keywords. 
%
According to the specification, the stream dictionary shall contain a \emph{Length} 
name to specify the length of the stream. In practice this can be omitted as 
long as the delimiting keywords are in place.
%
PDF supports encoding of streams, in which
case the dictionary describe which filters are required to decode the stream.

@@

% trailer
\subsubsection{Trailer}
The trailer is composed of a trailer dictionary, a pointer to the cross-reference 
section and an end-of-file marker. The trailer dictionary is introduced by the 
"trailer" keyword. 
%
\emph{Root} is in practice the only mandatory name in the trailer dictionary, referencing the root of the 
%
document structure. The "startxref" keyword is followed by the number of bytes from 
the beginning of the file to the first cross-reference section. As with the 
cross-reference section, it can be omitted. The string "\%\%EOF" marks the 
end-of-file, but can be omitted.

\subsection{PDF-based polyglots}
\label{sec:pdf-based-polyglots}
Several design choices in the PDF specification make the format particularly 
suitable for making polyglots. One such design choice is the error tolerant 
parser. This is in part motivated by PDF being a container format, 
%of which the 
another design choice which implies
%implication is 
that a PDF file can, by design, contain foreign syntax that 
could interfere with the syntax of the file. Another motivation is that PDF 
files are designed to be both forward- and backward-compatible. Readers 
implementing an older version of the specification do not recognize new 
features and behave as if they were not present in the file. 
%
The implementation notes of the specification describes some exceptions to the
requirements of the specification, such as the header can occur 
anywhere within the first 1024 bytes of the file. This flexibility gives plenty of room for 
combining with syntax of another format.
%
The specification declares many components to be required in a
PDF file, but as can be seen in Section~\ref{sec:format}, in practice several 
components can be omitted~\cite{Wolf:CCC10}. The code in Listing~\ref{lst:minimal-pdf} shows the 
minimal syntax required for a malformed, but valid PDF file.
%incrementally 
%updated by appending updated versions of existing objects to 
%the end of the file. In fact, 
%and formats. This and the fact that it is designed to be robust against errors,
%provides us with a format that is powerful, flexible and ideal for creating
%malicious polyglots.

@@

Furthermore the PDF format is of particular interest to us because of its 
many capabilities. Some examples include executing JavaScript, launching commands, 
and issuing HTTP requests. The HTTP requests are restricted to the origin of the 
PDF file, and will include any cookies associated with that origin. 
Adobe Reader also includes a flash 
runtime to play embedded flash files on systems that do
not have the Flash runtime installed. %Considering that it might be a conscious decision
%of the user to refuse a flash player,

@@

When a PDF document is embedded in a web page, the corresponding plugin 
is executed to render the content. Recall from Section~\ref{sec:cross-origin} 
that the plug-in is selected based 
on the MIME-type, either supplied by the server or in the type attribute
%
with preference for the later.
%
Also recall
that the browser rely on the plug-in to handle the situation when the MIME-type 
supplied by the server is inconsistent with the MIME-type of the plug-in.
In the case of the Adobe Reader plug-in, it disregards the MIME-type
supplied by the server and will attempt to interpret any content as PDF. 
Listing~\ref{lst:embed-pdf} shows how 
arbitrary content can be rendered as PDF format.

\begin{lstlisting}[label=lst:embed-pdf, caption=HTML for embedding PDF content]
<embed type="application/pdf" 
src="vulnerable.com/content?input=%PDF...">
\end{lstlisting}

@@

As the scenario details in Section~\ref{sec:injection}, the attacker in a syntax 
injection attack manipulates a vulnerable service to include PDF syntax in existing content, e.g. HTML-documents.
The PDF syntax is typically injected through user input used by the vulnerable service in the composition of documents.
%
An example of a suitable fragment to inject is shown in Listing~\ref{lst:xml-pdf}.
%
The content is subsequently 
embedded in the attackers page as PDF, as exemplified in Listing~\ref{lst:embed-pdf}.
As mentioned in Section~\ref{sec:pdf-based-polyglots}, the embedded PDF can issue 
requests to the origin it came from, carrying the cookies associated with that origin.
These requests allow for the extraction of sensitive user data that can either be communicated back to the attacker, or be leveraged in further attacks.

Thus far, exploitation of vulnerable services have been discussed, excluding the specific 
conditions under which a service is vulnerable. In order to exploit the 
vulnerability, the injected syntax must pass through any existing filters 
unchanged or at least with its semantics preserved. If user input in an 
HTML-document is not sanitized, any syntax would be unchanged and 
the service is vulnerable to less sophisticated attacks, e.g. cross-site scripting, 
therefore sanitized user input is of greater concern.
%
Based on the PDF samples in Listings~\ref{lst:minimal-pdf} and~\ref{lst:xml-pdf}, we can derive 
the set of tokens required to build a PDF. Assuming that alpha-numeric characters 
pass through a sanitization filter unmodified, the set of tokens is $\{\%, <<, >>, /\}$.
As can be noted, there is a small but significant overlap with the tokens of HTML, 
which implies that many websites protected against cross-site scripting attacks 
are also protected against PDF-based polyglot attacks. One of the problems of 
filtering input for inclusion in a web page are the many contexts in which the 
input can be included. A problem made even more complex by the diversity of 
languages the page contains. Language incompatibilities force context dependent 
filtering, where the same input is treated differently based on the context in 
which it is included. In certain contexts angle brackets are often left untouched by filters.

@@

\subsection{Content smuggling}
\todo{Deal with this section, extend and polish}
Due to the nature of the PDF, it can without much effort be combined with just 
about any other format. This provides ample opportunity to create malicious 
polyglots where PDF is either the benign or the malicious format. Consequently, 
this significantly expands the attack surface, making it important to take 
measures to protect against these attacks.

\paragraph{PDF as the benign format}
\todo{Services that accept PDFs, PDFs can be used to hide another malicious format}
Services like job brokers commonly let the user upload a CV in the form of a PDF-file.
Before such PDF-files are published to recruiters, they are verified to not 
contain any malicious payload. An attacker can produce a PDF that is valid and 
benign, but also a polyglot hiding another malicious format, such as Flash. 
A proof-of-concept can easily be created by storing the PDF as a static string 
variable in the malicious flash source code. When the flash is compiled, the 
source of the PDF is part of the output, making it a valid PDF.

@@

\subsection{PDF communication methods}
%\todo{rewrite section, focus on embedded flash, xxe and code samples}
%
In order to extract information and leak across origins, a communication channel
is required. PDF documents provide multiple ways of generating HTTP requests. 
Many of these are cross-origin, but, out of security concerns, only one-way in the sense that 
the PDF document will never see the result. Therefore, the focus is directed towards 
two methods that allow bidirectional communication. Since the document will retain the 
origin from which it was served, all 
requests issued from the document will include any cookies associated with the target origin.

%When embedded in a web page, many of the PDF communication methods utilize
%the underlying browser to make requests by navigating the browser to the target 
%URL. This will unload the current page in which the document is embedded. Arbitrary page 
%navigation is allowed with respect to SOP, since the response is not visible 
%to the document. Such methods can be used in CSRF attacks, but would 
%defeat the purpose of requesting information as the document cannot process the 
%response.

% Cross-origin leaks
%To leak information across origins, a communication channel is 
%required between the attacker and the target. PDF provides the necessary means 
%to mediate this information leakage. The flow of information can be divided into 
%two steps, extracting information from the target and delivering it to the attacker.
%Each requiring its own communication methods.

%\subsubsection{The target}
%There are two methods for a PDF file to communicate with the server it originates from, 
%both allowing it to read the response: from JavaScript using XML external entities 
%and from an embedded flash file.

\paragraph{XML External Entities}
The PDF JavaScript API includes a method to parse XML documents, called 
XMLData.parse. The XML document being processed may in turn rely on external entities, 
required for the parsing of the document, which the XML parser will request. The request is 
bound to the origin of the document, and the response is included in the XML document. 
The source code of the XML document reflects this response and can be retrieved at a later point. This 
results in a bidirectional communication channel.
%
As the response 
is included in the XML, the result has to be well-formed XML. This puts a 
restriction on the content that can be requested using this method. Considering 
that HTML pages are rarely well-formed this method may be too restricted in practice.

Listing~\ref{lst:xml-pdf} contains an example PDF document that uses XML external 
entities. Note that the one-way method \emph{getURL} is used to communicate the 
information back to the attacker. Given the compact size of the example, it is useful for the syntax 
injection scenario in Section~\ref{sec:injection}, which require injection of 
small syntax fragments.

\begin{lstlisting}[label=lst:xml-pdf, caption=PDF using XML for communication]
%PDF-
1 0 obj<<>>stream
xml = '<!DOCTYPE foo[<!ENTITY x SYSTEM "'+
	URL+'">]><foo>&x;</foo>'
var doc = XMLData.parse(xml);
getURL('attacker.com/?'+doc.saveXML())
endstream
trailer<</Root
   <</Pages<<>>
      /OpenAction
      <</S/JavaScript/JS 1 0 R>>
   >>
>>
\end{lstlisting}

\paragraph{Embedded Flash}
By embedding a Flash file in the PDF document the capabilities of the format can be 
extended even further. The Flash runtime supports bidirectional communication in 
accordance with SOP.
%
The embedded Flash inherits the origin of the PDF document, thus it can 
request and read any document from the originating server. Unlike the XML
method, there are no restrictions on what content can be requested, thus making 
it the most versatile of the available communication methods. The only drawback 
using this approach is the significant increase in terms of file size. Even the compact 
flash code in Listing~\ref{lst:flash-pdf} will result in a 6 kB Flash file. 
This suggests that this method might be better suited for the content smuggling 
scenario in Section~\ref{sec:smuggling}, rather than the syntax injection scenario.

\begin{lstlisting}[label=lst:flash-pdf, caption=Flash code for communication]
package {
  import flash.net.*;
  import flash.display.Sprite;
  public class Secret extends Sprite {
    var u:String;
    var r:URLRequest;
    var l:URLLoader;
    public function Secret() {
      u = 'vulnerable.com/secret';
      r = new URLRequest(u);
      l = new URLLoader(r);

      l.addEventListener('complete', 
        function():void {
          u = 'attacker.com/?' + r.data;
          r = new URLRequest(u);
          l = new URLLoader(r);
        });
    }
  }
}
\end{lstlisting}

Cross-origin communication in Flash adheres to the same-origin policy. However, this is not a restriction, since
 Flash also supports cross-origin communication via  
cross-origin resource sharing. Using CORS, a web server can relax the SOP to allow access to specified content.
%As mentioned, the embedded Flash file inherits the origin of the document.
An attacker can set an allow-all cross-origin policy, such as in Listing~\ref{lst:crossdomain} 
that open up for two way communication.

\begin{lstlisting}[label=lst:crossdomain, caption=Allow-all crossdomain.xml]
<?xml version="1.0"?>
<cross-domain-policy>
  <allow-access-from domain="*" />
</cross-domain-policy>
\end{lstlisting}

%\subsubsection{The attacker}
%There are a number of ways to communicate with the attacker, either directly 
%via the embedding page or through the attackers server. As mentioned previously, 
%many methods trigger page navigation in the browser. This will result in a 
%noticeable change in the victims browser. While it fulfills the purpose of 
%leaking information, an attacker can use more covert methods of communication
%to refrain from alerting the victim.

%\paragraph{Fragment identifier}
%The methods that navigate the browser can still be used for a somewhat covert 
%communication with the attacker. Navigating to the same page with a different 
%fragment identifier will not reload the page. Embedding the information in the fragment 
%identifier of the URL of the embedding website, creates a one-way communication 
%channel from the document to the page. Since the fragment identifier is visible 
%in the address field of the browser, an alert user would notice it changing.

%\paragraph{PostMessage}
%The PDF plugin includes an API for communicating directly with the page in 
%which it is embedded. The document can send messages to the page using 
%the hostContainer.postMessage method. The page registers a handler to receive 
%the event and is also able to post messages back to the document.
%This two way communication is not visible to the victim.

%\paragraph{Embedded Flash}

@@

\paragraph{Firefox, Safari and Opera}
The browsers, Firefox, Safari and Opera, are all susceptible to the attacks outlined as 
per Section~\ref{sec:vuln}, without any restrictions or modifications.

@@

\todo{What about serving as attachment, so that it is not viewed within the browser? How does it work with embedded content?}
\todo{Problem of transferring credentials between domains for services where the content requires authentication}

@@

Wolf's OMG WTF PDF presentation~\cite{Wolf:CCC10} is one of the 
inspirations for our work. The presentation explores the liberty
of the PDF format. In addition, it highlights that PDF interpreters
often disregard the specification demands. This is particularly
relevant as it allows crossbreeding PDF with such
formats as ZIP and EXE.
